Seasonality in governing submesoscale dynamics in the Kuroshio Extension

### 2016-07-04 writeup/rocha_etal.tex
Teriz last fixes.

% Shorter version of title entered in capital letters:
\titlerunninghead{Seasonality at submesoscales in the Kuroshio Extension}

@@

\title{Seasonality in governing dynamics at submesoscales in the Kuroshio Extension}
%
% e.g., \title{Terrestrial ring current:
% Origin, formation, and decay $\alpha\beta\Gamma\Delta$}
%

@@

\keypoints{\item  Upper-ocean submesoscale (10-100 km) turbulence and inertia-gravity waves
                  undergo strong seasonal cycles that are out-of-phase.
           \item  Submesoscale turbulence dominates the horizontal velocity and
                  sea-surface height variability in late Winter/early Spring.
            \item Inertia-gravity waves dominate the horizontal velocity and
                   sea-surface height variability in late Summer/early Fall.
                  }

@@

\begin{abstract}
    Two new high-resolution numerical simulations with embedded tides suggest a
    strong modulation in the governing near-surface dynamics at submesoscales
    (roughly 10-100 km) in the Kuroshio Extension. Consistent with recent studies, deep late-Winter mixed layers
    are prone to baroclinic instabilities, and submesoscale turbulence
    prevails in late Winter/early Spring. While summertime
    re-stratification weakens submesoscale turbulence, it also enhances inertia-gravity
     waves near the surface. In the Kuroshio Extension,
    inertia-gravity waves strongly dominate the submesoscale surface kinetic energy and
    sea-surface height variance in late Summer/early Fall.
\end{abstract}

@@

Using the output of $1/24^\circ$ and $1/48^\circ$ global
numerical simulations with embedded tides, we show that inertia-gravity waves undergo
a strong seasonality near the surface in the Kuroshio Extension region.
Interestingly, the seasonal cycle of inertia-gravity waves is out-of-phase
with the seasonal cycle of submesoscale turbulence. Consistent with previous studies,
deep late Winter mixed layers are prone to
shallow baroclinic instabilities that are roughly in geostrophic balance
and flux energy upscale \citep{sasaki_etal2014,callies_etal2016},
driving a mild seasonal modulation of the mesoscales \citep{sasaki_etal2014,qiu_etal2014}.

Inertia-gravity waves, on the other hand, peak in late Summer/early Fall,
when the upper ocean is strongly stratified. Thus there exists a
strong seasonal modulation in the dominant upper-ocean submesoscale dynamics:
submesoscale turbulence dominates the upper-ocean dynamics in late Winter/early Spring,
whereas quasi-linear inertia-gravity waves prevail in late Summer/early Fall.
Because submesoscale turbulence is weakest in late Summer/early Fall,
the present results suggest that inertia-gravity waves account for most of the
summertime submesoscale SSH variability.

@@

To study seasonal variations in the upper-ocean dynamics, we focus on the northwest
Pacific, in the vicinity of the Kuroshio
Extension, where previous studies have suggested strong submesoscale seasonality
\citep{sasaki_etal2014,qiu_etal2014}.
We analyze a sub-domain of the LLC4320 and LLC2160 simulations of about 2000 km$^2$
spanning 155-175$^\circ$E; 25-40$^\circ$N (figure \ref{fig1}a). The stratification
in this mesoscale-rich subtropical region undergoes a vigorous seasonal cycle: wintertime-enhanced
small-scale turbulence de-stratifies the upper ocean, yielding mixed layers
as deep as 300m. In late Spring/early Summer enhanced solar radiation and mixed-layer
instabilities re-stratify the upper ocean, yielding mixed layers as shallow as 40m.
Fundamentally, the upper-ocean density structure is well-captured by both LLC simulations:
a comparison with Argo climatology shows that both simulations skillfully represent the Kuroshio
Extension stratification and its seasonal variability  (supplemental material).

\begin{figure*}[ht]
\begin{center}
\hspace{-1.25cm}\includegraphics[width=.75\textwidth]{figs/fig1_1.pdf}\\
\vspace{-.125cm}
\includegraphics[width=.75\textwidth]{figs/fig1_2.pdf}
 \caption{(a) The study region with the subregion where the LLC outputs are
          analyzed. Colors represent the topography and white lines are contours of absolute
          dynamic topography every 0.1 m. LLC 4320 (1/48$^\circ$) snapshots of surface vorticity (b and c) and a transect
          of potential density at 165$^\circ$E (d and e). The snapshots were
          taken at 00:00 UTC.}
\vspace{-1.5cm}
 \label{fig1}
 \end{center}
 \end{figure*}

@@

Figures \ref{fig1}b-c show snapshots of vertical vorticity $\zeta$ in early Spring
(April 15) and Fall (October 15) in the LLC4320 (1/48$^\circ$) simulation.
The model solutions depict seasonality in vorticity: large values of
fine-grained vertical vorticity are observed in early Spring with maximum
values as large as $4f$, where $f$ is
the local planetary vorticity, and root-mean-square (RMS) of about $0.4f$. In early
Fall, the situation is the opposite: the vertical vorticity is relatively coarse-grained;
its local maximum and RMS are both smaller than $0.5f$.
Indeed, the seasonal cycle of vorticity and rate of strain are strong (figure~\ref{fig2}):
in both simulations, the RMS vorticity and strain are about twice as large in
late Winter/early Spring
than in late Summer/early Fall. Because the wintertime vorticity and strain rate
are dominated by the smallest scales in the flow (the KE spectra is shallower than
a $-3$ power-law in Winter), increasing the resolution from
1/24$^\circ$ to 1/48$^\circ$ significantly increases the wintertime RMS vorticity
and strain by about 40$\%$.

@@

The bulk of vertical vorticity and strain rate are associated with subinertial flows
($T_f = 2\pi/f_0\approx 23.5$h, where $f_0$ is the inertial frequency at
the mean latitude): daily-averaging the velocity fields suppresses superinertial
motions and reduces the RMS
vorticity by less than 40$\%$ and the RMS strain by less than 10$\%$; the seasonal
cycle remains strong (see
dashed lines in figures \ref{fig2}a-b). Most of this seasonal cycle is associated
with submesoscale flows: smoothing the velocity fields
with a Hanning filter with cut-off scale of 100 km dramatically reduces the RMS
vorticity and strain. The reduction in variance is about 80$\%$ in winter,
yielding RMS vorticity and strain roughly consistent with the diagnostics from
AVISO gridded geostrophic velocities (compare red lines to black lines in figures \ref{fig2}a-b).
The picture that emerges is consistent with recent studies: shallow baroclinic
instabilities energize the submesoscales in late Winter, drawing from the available
potential energy stored in deep mixed layers \citep{sasaki_etal2014,callies_etal2015,callies_etal2016}.

The seasonal cycle of the horizontal divergence, however, showcases the complexity
of the upper-ocean annual variability. If submesoscale eddies and fronts dominated
the near-surface variability all year, then the seasonal cycle of horizontal divergence,
vertical vorticity, and lateral strain rate would be in phase \citep[e.g.,][]{sasaki_etal2014}.
While there is a clear wintertime peak in divergence of daily-averaged velocity
(see dashed lines in figure \ref{fig2}c; RMS divergence $\sim0.22 f$ in the 1/48$^\circ$
simulation), the hourly fields show
a stronger enhancement of
lateral divergence in late Summer/early Fall (RMS divergence $\sim0.22 f$ the 1/48$^\circ$
simulation). Because the 1/48$^\circ$ simulation better resolves
smaller-scale submesoscale flows, a secondary RMS divergence peak in winter  is
nearly as strong as in Summer. Submesoscale fronts and eddies evolve
relatively fast, and there is no clear
temporal and spatial scale separation between those motions and inertia-gravity waves
\citep{mcwilliams2016}:
daily-averaging the velocity fields efficiently suppresses the summertime horizontally
divergent flows, but also reduces the wintertime lateral divergence by about 50$\%$.
The vorticity and strain rate (Figure \ref{fig2}a-b) show that most of the lateral divergence is associated
with submesoscale flows: smoothing the velocity fields with a 100km-cut-off
suppresses more than 80$\%$ of the RMS divergence.

\section{Seasonality in governing submesoscale dynamics}
The overall results of figure \ref{fig2}c suggest that submesoscale surface
 variability stems from different dynamics in Summer than Winter.
 To characterize these differences, we calculate
joint probability distributions (jPDF) of
vorticity-strain and vorticity-Laplacian of sea-surface height (\ref{fig3}).

The April vorticity-strain jPDF has a shape characteristic of submesoscale turbulence
 (see figures \ref{fig3}a-c). The alignment of vorticity and strain $\alpha \sim \pm\zeta$
 with strong positive skewness are fingerprints of submesoscale fronts
 \citep{shcherbina_etal2013,mcwilliams2016}. The shape of the vorticity-strain
 jPDF is similar for hourly, daily-averaged, and residual velocities, although
 the vorticity skewness reduces from 1.4 to 1.13 from hourly to daily-averaged.
 The April results are characteristic of Winter, indicating that wintertime submesoscale
 surface velocity is strongly dominated by submesoscale turbulence --- this is true even for
 superinertial submesoscale currents. The hourly velocity fields are significantly
ageostrophic as depicted by the jPDF of vorticity-Laplacian of SSH
(figure \ref{fig3}g). Even in April, only the daily-averaged fields
are largely
in  geostrophic balance (the jPDF of daily-averaged vorticity vs. Laplacian of SSH
is an ellipse with large eccentricity and main axis tilted by 45$^\circ$; figure \ref{fig3}h).

@@

The October vorticity-strain jPDF shows much weaker skewness (the vorticity skewness
is 0.68 and 0.67 for hourly and daily-averaged velocities). The shape of the
vorticity-strain jPDF  appears to be a combination of two half-ellipses centered
about $\zeta=0$, one with a 45$^\circ$ slope (characteristic of submesocale fronts
that persist in Summer) and one with a very steep slope.
That the submesoscale dynamics in October are mainly unbalanced is clearly depicted
in the shape of jPDF of vorticity-Laplacian of SSH for hourly and
residual fields
(figure \ref{fig3}j and \ref{fig3}k), which is an ellipse aligned in the vertical axis.
Of course daily-averaging the model suppresses the ageostrophic flows, and
the daily-averaged flow is essentially geostrophic as depicted by the 45$^\circ$-tilted
ellipse in the vorticity-Laplacian of SSH jPDF.

 Time series of vorticity and divergence PDFs (supplemental material) suggest
 a strong oscillation between these two regimes. In late Winter/early Spring
 the vorticity is strongly positively skewed, whereas the divergence is moderately
 negatively skewed as predicted by frontogenesis \citep{capet_etal2008a,mcwilliams2016}
. In late Summer/early Fall, the divergence is stronger, but PDFs are much less skewed,
consistent with linear inertia-gravity waves.

@@

\section{Summary and Conclusion}
Our main finding is that in global simulations with embedded tides the
near-surface upper-ocean inertia-gravity waves in the Kuroshio Extension undergo
a strong seasonal cycle that is out-of-phase with the seasonal cycle of
submesoscale turbulence. \cite{dasaro1978} showed that velocity of linear internal waves
in the mixed-layer strongly depends on the density jump at the mixed-layer
base, with largest mixed layer velocities when the jump is strongest.
In Summer, the shallow mixed layer overlays a strong seasonal pycnocline,
and thus the internal waves projection onto the mixed-layer may be stronger
according to \cite{dasaro1978}'s arguments. An
alternative explanation is that the shape of the baroclinic modes
changes seasonally. In particular the baroclinic modes are
significantly more surface intensified in late Summer/early Fall
 (supplemental material).
If the internal wave source has weak seasonal dependence, then a strong
seasonal cycle in the near-surface expression of internal waves is expected.
Internal tide generation does not show a seasonal modulation \citep[e.g.,][]{alford2003},
and therefore it is plausible to expect a strong seasonality in the near-surface
expression of internal tides and
other small-scale internal waves generated through internal tide interactions.
Near-inertial wave generation peaks in Winter, but those waves project on large
horizontal scales \citep[e.g, ][]{qi_etal1995}.

@@

% That the  surface velocity and SSH submesoscale variability may be
% dominated by ageostrophic flows in
% Summer/Fall has consequences for the interpretation of data from
% the future SWOT and COMPIRA altimeter missions,
% which will deliver SSH measurements at submesoscales. To the extent that
% high-frequency flows are dominated by spatially incoherent internal tides and other
% internal waves, it may be very difficult (if not impossible) to
% separate SSH submesoscale variability associated with geostrophic motions
% from high-frequency, ageostrophic flows. Thus, previous claims that
% one will be able to easily obtain submesoscale surface
% geostrophic velocities and monitor such seasonal cycle on global scales
% \citep{sasaki_etal2014,qiu_etal2014} are overstated.

@@

% \section{Projection on time scales}
% To better understand the high-frequency/high-wavenumber flows  we calculate frequency
% spectra of surface KE (figure \ref{fig4}). To obtain statistically meaningful results,
% we average spectral estimates along a section at 165$^\circ$E.  There is about 50
% independent spectral realizations.
%
% The surface KE spectra of subinertial flows very roughly follow a $-2$ power-law.
% The high-frequency flows are split near-inertial flows, tidal signals and their
% higher harmonics. Hence these high-frequency flows are dominated by inertia-gravity
% waves. (Note that the inertial peak nearly merges with the
% diurnal tide peak.) There is statistically significant seasonality. The seasonal
% cycle of the sub-inertial motions (at least for periods between 5-20 days)
% is out-of-phase with the seasonal cycle of supra-inertial flows: the former flows
% are more energetic in April and the latter in October.
%
% Most of these high-frequency inertia-gravity waves project onto scales smaller
% than 100 km. The surface KE spectra of 100-km-smoothed velocity field is significantly
% suppressed in supra-inertial frequencies --- there are no statistically significant
% changes at sub-inertial frequencies (see dashed lines in figure \ref{fig4}).

@@

\newpage
\section*{Supplemental material 1: On the LLC simulations}

Table \ref{tab:llc} show details of the LLC spin-up hierarchy.
Both simulation analyzed in the present paper, LLC2160 and LLC4320,
were forced by surface fluxes are from the 0.14$^\circ$ European Centre for
Medium-range Weather Forecasting (ECMWF) atmospheric operational model analysis,
 starting 2011 and the 16 most significant  tidal components.
The LLC control files are available online
(at \texttt{http://mitgcm.org/viewvc/MITgcm/MITgcm$\_$contrib/
llc$\_$hires}).

@@

@@

\subsection*{Seasonality in statistics: probability density functions}
\begin{figure*}[ht]
   \begin{center}
     \includegraphics[width=1.\textwidth]{figs/fig_s2_2.pdf}
  \caption{Time dependence of probability density functions of vertical vorticity (a, b)
  and horizontal divergence (c, d). Monthly averages in April and October are
  also shown (e, f).}
  \label{figS2_2}
  \end{center}
\end{figure*}

@@

Figure \ref{figS2_1} shows a comparison between potential density sections
at 165$^\circ$E from
the Roemmich-Gilson Argo Climatology \citep[updated from][]{roemmich_gilson2009}.
and the LLC2160 and LLC4320 simulations. Of course, the Argo climatology is smoother
owing to lower resolution and long-term averaging. Nonetheless, simulations reproduce well the strong
seasonality in the stratification. The upper ocean is weakly stratified in late Winter/early Fall
and restratifies in Summer. Argo climatology and LLC simulations mixed-layer depths are consistent
at all seasons (not shown).

@@

@@

\section*{Supplemental material 3: Seasonality in pressure modes near-surface amplitude}

Figure \ref{figS3}a shows the domain-averaged potential density from the WOA 2013
climatology in April and October. Clearly, the changes to the stratification
are dramatic, with the formation of a seasonal pycnocline in Summer. The strong
seasonality of upper-ocean stratification yields a strong seasonality in the
near-surface shape and amplitude of baroclinic pressure modes (\ref{figS3}b).
In particular, the surface amplitude is much large in Summer. Thus the projection
of baroclinic tides on the surface may vary significantly seasonally.

\textcolor{red}{Perhaps include analytical results from a model with a slab over a slowly varying
pycnocline.}

>>>>

% Shorter version of title entered in capital letters:
\titlerunninghead{Seasonality at submesoscales}

@@

\title{Seasonality in governing submesoscale dynamics in the Kuroshio Extension}
%
% e.g., \title{Terrestrial ring current:
% Origin, formation, and decay $\alpha\beta\Gamma\Delta$}
%

@@

\keypoints{\item  Upper-ocean submesoscale (10-100 km) turbulence and inertia-gravity waves
                  undergo strong seasonal cycles that are out-of-phase.
           \item  Submesoscale turbulence dominates the horizontal velocity and
                  sea-surface height variability in late winter/early spring.
            \item Inertia-gravity waves dominate the horizontal velocity and
                   sea-surface height variability in late summer/early fall.
                  }

@@

\begin{abstract}
    Two new high-resolution numerical simulations with embedded tides suggest a
    strong modulation in the governing near-surface dynamics at submesoscales
    (roughly 10-100 km) in the Kuroshio Extension. Consistent with recent studies, deep late-winter mixed layers
    are prone to baroclinic instabilities, and submesoscale turbulence
    prevails in late winter/early spring. While summertime
    re-stratification weakens submesoscale turbulence, it also enhances inertia-gravity
     waves near the surface. In the Kuroshio Extension,
    inertia-gravity waves strongly dominate the submesoscale surface kinetic energy and
    sea-surface height variance in late summer/early fall.
\end{abstract}

@@

Using the output of $1/24^\circ$ and $1/48^\circ$ global
numerical simulations with embedded tides, we show that inertia-gravity waves undergo
a strong seasonality near the surface in the Kuroshio Extension region.
Interestingly, the seasonal cycle of inertia-gravity waves is out-of-phase
with the seasonal cycle of submesoscale turbulence. Consistent with previous studies,
deep late-winter mixed layers are prone to
shallow baroclinic instabilities that are roughly in geostrophic balance
and flux energy upscale \citep{sasaki_etal2014,callies_etal2016},
driving a mild seasonal modulation of the mesoscales \citep{sasaki_etal2014,qiu_etal2014}.

Inertia-gravity waves, on the other hand, peak in late summer/early fall,
when the upper ocean is strongly stratified. Thus there exists a
strong seasonal modulation in the dominant upper-ocean submesoscale dynamics:
submesoscale turbulence dominates the upper-ocean dynamics in late winter/early ,
whereas quasi-linear inertia-gravity waves prevail in late summer/early fall.
Because submesoscale turbulence is weakest in late summer/early fall,
the present results suggest that inertia-gravity waves account for most of the
summertime submesoscale SSH variability.

@@

To study seasonal variations in the upper-ocean dynamics, we focus on the northwest
Pacific, in the vicinity of the Kuroshio
Extension, where previous studies have suggested strong submesoscale seasonality
\citep{sasaki_etal2014,qiu_etal2014}.
We analyze a sub-domain of the LLC4320 and LLC2160 simulations of about 2000 km$^2$
spanning 155-175$^\circ$E; 25-40$^\circ$N winter \ref{fig1}a). The stratification
in this mesoscale-rich subtropical region undergoes a vigorous seasonal cycle: wintertime-enhanced
small-scale turbulence de-stratifies the upper ocean, yielding mixed layers
as deep as 300m. In late spring/early summer enhanced solar radiation and mixed-layer
instabilities re-stratify the upper ocean, yielding mixed layers as shallow as 40m.
Fundamentally, the upper-ocean density structure is well-captured by both LLC simulations:
a comparison with Argo climatology shows that both simulations skillfully represent the Kuroshio
Extension stratification and its seasonal variability  (supplemental material).

\begin{figure*}[ht]
\begin{center}
\hspace{-1.25cm}\includegraphics[width=.75\textwidth]{figs/fig1_1.pdf}\\
\vspace{-.125cm}
\includegraphics[width=.75\textwidth]{figs/fig1_2.pdf}
 \caption{(a) The study region with the subregion where the LLC outputs are
          analyzed. Colors represent the topography and white lines are contours of absolute
          dynamic topography every 0.1 m from AVISO. LLC 4320 (1/48$^\circ$) snapshots of surface vorticity (b and c) and transects
          of potential density at 165$^\circ$E (d and e). The snapshots were
          taken at 00:00 UTC.}
\vspace{-1.5cm}
 \label{fig1}
 \end{center}
 \end{figure*}

@@

Figures \ref{fig1}b-c show snapshots of vertical vorticity $\zeta$ in early spring
(April 15) and fall (October 15) in the LLC4320 (1/48$^\circ$) simulation.
The model solutions depict seasonality in vorticity: large values of
fine-grained vertical vorticity are observed in early spring with maximum
values as large as $4f$, where $f$ is
the local planetary vorticity, and root-mean-square (RMS) of about $0.4f$. In early
fall, the situation is the opposite: the vertical vorticity is relatively coarse-grained;
its local maximum and RMS are both smaller than $0.5f$.
Indeed, the seasonal cycle of vorticity and rate of strain are strong winter~\ref{fig2}):
in both simulations, the RMS vorticity and strain are about twice as large in
late winter/early spring
than in late summer/early fall. Because the wintertime vorticity and strain rate
are dominated by the smallest scales in the flow (the KE spectra is shallower than
a $-3$ power-law in winter), increasing the resolution from
1/24$^\circ$ to 1/48$^\circ$ significantly increases the wintertime RMS vorticity
and strain by about 40$\%$.

@@

The bulk of vertical vorticity and strain rate are associated with subinertial flows
($T_f = 2\pi/f_0\approx 23.5$h, where $f_0$ is the inertial frequency at
the mean latitude): daily-averaging the velocity fields suppresses superinertial
motions and reduces the RMS
vorticity by less than 40$\%$ and the RMS strain by less than 10$\%$; the seasonal
cycle remains strong (see
dashed lines in figures \ref{fig2}a-b). Most of this seasonal cycle is associated
with submesoscale flows: smoothing the velocity fields
with a Hanning filter with cut-off scale of 100 km dramatically reduces the RMS
vorticity and strain. The reduction in variance is about 80$\%$ in winter,
yielding RMS vorticity and strain roughly consistent with the diagnostics from
AVISO gridded geostrophic velocities (compare red lines to black lines in figures \ref{fig2}a-b).
The picture that emerges is consistent with recent studies: shallow baroclinic
instabilities energize the submesoscales in late winter, drawing from the available
potential energy stored in deep mixed layers \citep{sasaki_etal2014,callies_etal2015,callies_etal2016}.

The seasonal cycle of the horizontal divergence, however, showcases the complexity
of the upper-ocean annual variability. If submesoscale eddies and fronts dominated
the near-surface variability all year, then the seasonal cycle of horizontal divergence,
vertical vorticity, and lateral strain rate would be in phase \citep[e.g.,][]{sasaki_etal2014}.
While there is a clear wintertime peak in divergence of daily-averaged velocity
(see dashed lines in figure \ref{fig2}c; RMS divergence $\sim0.22 f$ in the 1/48$^\circ$
simulation), the hourly fields show
a stronger enhancement of
lateral divergence in late summer/early fall (RMS divergence $\sim0.22 f$ the 1/48$^\circ$
simulation). Because the 1/48$^\circ$ simulation better resolves
smaller-scale submesoscale flows, a secondary RMS divergence peak in winter  is
nearly as strong as in summer. Submesoscale fronts and eddies evolve
relatively fast, and there is no clear
temporal and spatial scale separation between those motions and inertia-gravity waves
\citep{mcwilliams2016}:
daily-averaging the velocity fields efficiently suppresses the summertime horizontally
divergent flows, but also reduces the wintertime lateral divergence by about 50$\%$.
The vorticity and strain rate winter \ref{fig2}a-b) show that most of the lateral divergence is associated
with submesoscale flows: smoothing the velocity fields with a 100km-cut-off
suppresses more than 80$\%$ of the RMS divergence.

\section{Seasonality in governing submesoscale dynamics}
The overall results of figure \ref{fig2}c suggest that submesoscale surface
 variability stems from different dynamics in summer than winter.
 To characterize these differences, we calculate
joint probability distributions (jPDF) of
vorticity-strain and vorticity-Laplacian of sea-surface height (\ref{fig3}).

The April vorticity-strain jPDF has a shape characteristic of submesoscale turbulence
 (see figures \ref{fig3}a-c). The alignment of vorticity and strain $\alpha \sim \pm\zeta$
 with strong positive skewness are fingerprints of submesoscale fronts
 \citep{shcherbina_etal2013,mcwilliams2016}. The shape of the vorticity-strain
 jPDF is similar for hourly, daily-averaged, and residual velocities, although
 the vorticity skewness reduces from 1.4 to 1.13 from hourly to daily-averaged.
 The April results are characteristic of winter, indicating that wintertime submesoscale
 surface velocity is strongly dominated by submesoscale turbulence --- this is true even for
 superinertial submesoscale currents. The hourly velocity fields are significantly
ageostrophic as depicted by the jPDF of vorticity-Laplacian of SSH
winter \ref{fig3}g). Even in April, only the daily-averaged fields
are largely
in  geostrophic balance (the jPDF of daily-averaged vorticity vs. Laplacian of SSH
is an ellipse with large eccentricity and main axis tilted by 45$^\circ$; figure \ref{fig3}h).

@@

The October vorticity-strain jPDF shows much weaker skewness (the vorticity skewness
is 0.68 and 0.67 for hourly and daily-averaged velocities). The shape of the
vorticity-strain jPDF  appears to be a combination of two half-ellipses centered
about $\zeta=0$, one with a 45$^\circ$ slope (characteristic of submesocale fronts
that persist in summer) and one with a very steep slope.
That the submesoscale dynamics in October are mainly unbalanced is clearly depicted
in the shape of jPDF of vorticity-Laplacian of SSH for hourly and
residual fields
winter \ref{fig3}j and \ref{fig3}k), which is an ellipse aligned in the vertical axis.
Of course daily-averaging the model suppresses the ageostrophic flows, and
the daily-averaged flow is essentially geostrophic as depicted by the 45$^\circ$-tilted
ellipse in the vorticity-Laplacian of SSH jPDF.

 Time series of vorticity and divergence PDFs (supplemental material) suggest
 a strong oscillation between these two regimes. In late winter/early spring
 the vorticity is strongly positively skewed, whereas the divergence is moderately
 negatively skewed as predicted by frontogenesis \citep{capet_etal2008a,mcwilliams2016}
. In late summer/early fall, the divergence is stronger, but PDFs are much less skewed,
consistent with linear inertia-gravity waves.

@@

\section{Summary and Conclusion}
Our main finding is that in global simulations with embedded tides the
near-surface inertia-gravity waves in the Kuroshio Extension undergo
a strong seasonal cycle that is out-of-phase with the seasonal cycle of
submesoscale turbulence. \cite{dasaro1978} showed that velocity of linear internal waves
in the mixed-layer strongly depends on the density jump at the mixed-layer
base, with largest mixed layer velocities when the jump is strongest.
In summer, the shallow mixed layer overlays a strong seasonal pycnocline,
and thus the internal waves projection onto the mixed-layer may be stronger
according to \cite{dasaro1978}'s arguments. An
alternative explanation is that the shape of the baroclinic modes
changes seasonally. In particular the baroclinic modes are
significantly more surface intensified in late summer/early fall
 (supplemental material).
If the internal wave source has weak seasonal dependence, then a strong
seasonal cycle in the near-surface expression of internal waves is expected.
Internal tide generation does not show a seasonal modulation \citep[e.g.,][]{alford2003},
and therefore it is plausible to expect a strong seasonality in the near-surface
expression of internal tides and
other small-scale internal waves generated through internal tide interactions.
Near-inertial wave generation peaks in winter, but those waves project on large
horizontal scales \citep[e.g, ][]{qi_etal1995}.

@@

% That the  surface velocity and SSH submesoscale variability may be
% dominated by ageostrophic flows in
% summer/fall has consequences for the interpretation of data from
% the future SWOT and COMPIRA altimeter missions,
% which will deliver SSH measurements at submesoscales. To the extent that
% high-frequency flows are dominated by spatially incoherent internal tides and other
% internal waves, it may be very difficult (if not impossible) to
% separate SSH submesoscale variability associated with geostrophic motions
% from high-frequency, ageostrophic flows. Thus, previous claims that
% one will be able to easily obtain submesoscale surface
% geostrophic velocities and monitor such seasonal cycle on global scales
% \citep{sasaki_etal2014,qiu_etal2014} are overstated.

@@

% \section{Projection on time scales}
% To better understand the high-frequency/high-wavenumber flows  we calculate frequency
% spectra of surface KE winter \ref{fig4}). To obtain statistically meaningful results,
% we average spectral estimates along a section at 165$^\circ$E.  There is about 50
% independent spectral realizations.
%
% The surface KE spectra of subinertial flows very roughly follow a $-2$ power-law.
% The high-frequency flows are split near-inertial flows, tidal signals and their
% higher harmonics. Hence these high-frequency flows are dominated by inertia-gravity
% waves. (Note that the inertial peak nearly merges with the
% diurnal tide peak.) There is statistically significant seasonality. The seasonal
% cycle of the sub-inertial motions (at least for periods between 5-20 days)
% is out-of-phase with the seasonal cycle of supra-inertial flows: the former flows
% are more energetic in April and the latter in October.
%
% Most of these high-frequency inertia-gravity waves project onto scales smaller
% than 100 km. The surface KE spectra of 100-km-smoothed velocity field is significantly
% suppressed in supra-inertial frequencies --- there are no statistically significant
% changes at sub-inertial frequencies (see dashed lines in figure \ref{fig4}).

@@

\newpage
\section*{Details of the LLC simulations}

Table \ref{tab:llc} shows details of the LLC spin-up hierarchy.
Both simulations analyzed in the present paper, LLC2160 and LLC4320,
were forced by surface fluxes from the 0.14$^\circ$ European Centre for
Medium-range Weather Forecasting (ECMWF) atmospheric operational model analysis,
 starting in 2011, and the 16 most significant  tidal components.
The LLC control files are available online
(at \texttt{http://mitgcm.org/viewvc/MITgcm/MITgcm$\_$contrib/
llc$\_$hires}).

@@

The spin-up of the upper-ocean in the LLC4320 (1/48$^\circ$) simulation is
a couple of days, as depicted the time series of surface horizontal divergence
in Figure 2b of the main text (see initial rapid increase in RMS divergence in
September 2011). To avoid the spin-up stage, we chose October to characterize the
late summer/early fall. The late winter/early spring waves chosen to be 6 months
out-of-phase with late summer/early fall (hence April).

@@

\subsection*{Seasonality in probability density functions}

Probability density functions (pdfs) of vertical vorticity and horizontal divergence
 at the surface (Figure \ref{figS2_2}) depicts the strong seasonality in the
 submesoscale dynamics. Time series of vorticity pdf indicate that
 the development of positive skewness in winter (Figures \ref{figS2_2}a-b) both in
 hourly ($\sim 1.4$) and daily-averaged ($\sim1.1$) fields. The skewness
 in divergence for hourly fields is very small year-round ($\sim -0.24$), suggesting the
 prevalence of inertia-gravity waves. Filtering out the waves by daily-averaging
 the velocity field increases the skewness in divergence, particularly in winter ($\sim-0.64$). Both
positive vorticity (cyclonic) skewness and negative divergence (convergence) skewness
are characteristics of submesoscale turbulence \citep{mcwilliams2016}.

\begin{figure*}[ht]
   \begin{center}
     \includegraphics[width=1.\textwidth]{figs/fig_s2_2.pdf}
  \caption{Time dependence of probability density functions of vertical vorticity (a, b)
  and horizontal divergence (c, d). Monthly averages in April and October are
  also shown (e, f).}
  \label{figS2_2}
  \end{center}
\end{figure*}

@@

Figure \ref{figS2_1} shows a comparison between potential density sections
at 165$^\circ$E from
the Roemmich-Gilson Argo Climatology \citep[updated from][]{roemmich_gilson2009}.
and the LLC2160 and LLC4320 simulations. Of course, the Argo climatology is smoother
owing to lower resolution and long-term averaging. Nonetheless, simulations reproduce well the strong
seasonality in the stratification. The upper ocean is weakly stratified in late winter/early fall
and restratifies in summer. Argo climatology and LLC simulations mixed-layer depths are consistent
at all seasons (not shown).

@@

\subsection*{Depth-dependence}
Figure \ref{figS2_2} shows the depth-dependence of the  root-mean-square of the velocity
gradient lateral tensor for the LLC4320 simulation. The strong seasonality, as depicted
at the surface in Figure 2 of the main manuscript, is confined to the mixed layer, which
varies from 50 m in summer to about 350 m in winter.

\begin{figure*}[ht]
   \begin{center}
     \includegraphics[width=.9\textwidth]{figs/fig_s2_2.png}
  \caption{Depth-dependence of root-mean-square of vertical vorticity (a,d), rate of lateral strain (b,e)
  and horizontal divergence (c,f) for hourly (a,b,c) and daily-averaged (d,e,f,) fields.}
  \label{figS2_2}
  \end{center}
\end{figure*}

@@

\section*{Seasonality in pressure modes near-surface amplitude}

Figure \ref{figS3}a shows the domain-averaged potential density from the WOA 2013
climatology in April and October. Clearly, the changes to the stratification
are dramatic, with the formation of a seasonal pycnocline in summer. The strong
seasonality of upper-ocean stratification yields a strong seasonality in the
near-surface shape and amplitude of baroclinic pressure modes (\ref{figS3}b).
In particular, the surface amplitude is much large in summer. Thus the projection
of baroclinic tides on the surface may vary significantly seasonally.

%\textcolor{red}{Perhaps include analytical results from a model with a slab over a slowly varying
%pycnocline.}
